---
title: "Budget Data Cleaning"
author: "Chelsea Ursaner & Brett Kobold"
date: "5/18/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Incremental changes

Changes from previous years with descriptions

```{r}
library(dplyr) #moving and shaping the data
library(tidyr) #allows for melt and cast
library(magrittr) #pipeforwarding package %>% 
library(readxl) #reading the excel file
library(stringr) #working with strings eaiser
library(RSocrata)

#Read in data from data.lacity.org and downloaded files
#Pull existing data from Socrata
inc_existing <- read.socrata(url = 'https://data.lacity.org/A-Prosperous-City/General-City-Budget-Incremental-Changes/k4k6-bwwv')
#transfer old dataset to tidy format
inc_existing_tidy <- gather(inc_existing, "Budget", "n", 16:17, na.rm=T)

#Read in data from new excel file
setwd("~/R_labs/laBudget/FY17-18 Adopted Budget")
inc_new <- read_excel("Budget Requests Detail Sec2_1718Adopted.xls", 1) %>% data.frame()
#transfer new dataset to tidy format
inc_new_tidy <- gather(inc_new, "Budget", "n", 14:15, na.rm=T)

###Make master dataset
#Make sure column names match
inc_existing_tidy <- rename(inc_existing_tidy, One.Time..On.going = One.Time..01....On.going..BB.)
inc_new_tidy <- rename(inc_new_tidy, Account.Code = Budget.Object.Code, Account.Name = Audit.Budget.Object.Name)

inc <- rbind(inc_existing_tidy, inc_new_tidy)
inc <- inc %>% rename(Incremental.Change = n)

write.csv(inc, "incremental.csv", row.names=F)

```

##Appropriations

Appropriations are also known as budget expenses or allocations. The challenging piece here is to assign the 'priority outcome' which can fall into four categories: Make LA the Best Run City in America, Make LA the Safest.., Provide Good Jobs to All Angelenos, and Make LA the Most Livable and Sustainable Big City..

```{r}
library(dplyr) #moving and shaping the data
library(tidyr) #allows for melt and cast
library(magrittr) #pipeforwarding package %>% 
library(readxl) #reading the excel file
library(stringr) #working with strings eaiser
library(RSocrata)

#reading in data from data.lacity.org and downloaded files
#Pull existing data from Socrata
appropriations_existing <- read.socrata(url = 'https://data.lacity.org/A-Prosperous-City/Open-Budget-Appropriations-Fiscal-Years-2010-2018/5242-pnmt')


###Remove proposed to replace with adopted
appropriations_existing <- appropriations_existing %>% filter(Fiscal_Year != 2018) 
#Export to start working in Excel to do a vlookup
write.csv(appropriations_existing, "working_appropriations.csv", row.names = F)

#Pull new data from Excel extract
setwd("~/R_labs/laBudget/FY17-18 Adopted Budget")
appropriations_new <- read_excel("Expenditures_Sec2 All Regular Depts and NonDepts_1718Adopted.xls", 1) %>% data.frame()

#make it look like the original dataset
appropriations_new <- appropriations_new %>% select(-c(13:16))

#Export to start working in Excel to do a vlookup
#working dataset
write.csv(appropriations_existing, "working_appropriations.csv", row.names = F)
#new values
write.csv(appropriations_new, "new_appropriations.csv", row.names = F)
```

... and do the rest in excel, tableau, and datasync unless we want to try it with R

```{r}
#Read the new new appropriations file
#setwd("~/R_labs/laBudget/FY17-18 Adopted Budget")
appropriations_new <- read.csv("new_new_appropriations.csv")

#Make sure they match
names(appropriations_existing)
names(appropriations_new)

appropriations = rbind(appropriations_existing, appropriations_new)
write.csv(appropriations, 'appropriations_final.csv', row.names = F)
#replace using datasync
```
